---
title: "ARIMA Coding"
author: "Steve Midway"
date: "Spring 2020"
output:
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As with most things in R, there is a rich environments of packages and functions beyond the basics presented here. This coding tutorial is not intended to be comprehensive, but to simply serve as a way to get you coding ARIMA models and understanding their mechanics. This document is intended for you to work through. Much of the code is there, but much of it will need to be built on. 

### Libraries
```{r}
library(astsa) # Applied Statistical Time Series Analysis
```

### Time Series Objects
When working with time series data in R, we typically want to have the observations in a time series object, or `ts()`.

```{r, echo=F, eval=F}
Review ts objects
apply ts() function to convert to ts class
plot ts object
can specify start date, frequency,

play with simulating different combinations of ar and ma and area
arima.sim

…then use arima model to estimate the simulated data (or use sarima)

Example:

AR only
# Plot x
x <- arima.sim(model = list(order = c(2, 0, 0), ar = c(1.5, -.75)), n = 200)
plot(x)

# Plot the sample P/ACF of x
acf2(x)

# Fit an AR(2) to the data and examine the t-table
sarima(x, 2, 0, 0)


MA only
# Plot x
x <- arima.sim(model = list(order = c(0, 0, 1), ma = -.8), n = 100)
plot(x)

# Plot the sample P/ACF of x
acf2(x)

# Fit an MA(1) to the data and examine the t-table
sarima(x, 0,0,1)

sarima.for()

```

